---
title: "RLTS Contacts"
---

R Package from the Lanzas Lab: https://github.com/lanzaslab/contact
```{r}
# devtools::install_github("lanzaslab/contact")
library(contact)
library(dplyr)
library(sf)
```


Here is an example of how to read all the sheets in a data frame from a *xlsx* file.
```{r eval = F}
L <- list()
for(g in 2:5){
  # g <- 5
  L[[g]] <- readxl::read_xlsx(path = "../../Box Sync/MPVM/RP/PERU 2016 DATA/Vacunacion/Data/VacDepDisProv_2016.xlsx", sheet = g)
  L[[g]]$species <- i
}



DF <- do.call(rbind, L)
```



# Create a point-location-based environmental contact network

The network will represent contacts between calves in a single feedlot pen with theut water source. We will identify the number of "contacts" each individual had with the water sourcein the pen. A "contact" is defined as occuring when point-locations were within a pre-determine spatial-treshold distance (SpTh) of the water source.  We wil set a SpTh of 0.333 m  (i.e. the approximate distance from RFID tags to calves noses), then re-define this SpTh to account for RTLS accuracy using the `findDistTresh` function.  

```{r Get the data}
data("calves")
calves %>%
  head()
```


## Ensure all required columns exists in the dataset.

```{r}
# we must append the columns date and time together first:
calves <- calves %>% 
  mutate(DateTime = paste0(date, ":", time)) %>% # Paste the date and time volumns
  mutate(DateTime = as.POSIXct(DateTime, format = "%m-%d-%Y:%H:%M:%S")) # Format as date time (POSIXct)
```

## Calculate the distances between the water polygon and the calves at each time step
```{r}
waterSp <- cbind(x = c(61.43315, 61.89377, 62.37518, 61.82622, 61.43315), y = c(62.44815, 62.73341, 61.93864, 61.67411, 62.44815)) # this is the location of the water source
waterSp <- st_polygon(list(waterSp)) # convert to a spatial polygon
# convert the calves location to spatial points
calvesSp <- calves %>%
  st_as_sf(coords = c("x", "y")) 

# Visualize the data:
# Plot the calves
calvesSp %>%
  st_geometry() %>%
  plot(pch = 16, cex= 0.5)
# Plot the water source
waterSp %>%
  st_geometry() %>%
  plot(add = T, col = "red")
```


```{r Get the distance}
# Calculate the distance to water source
calvesSp$WaterDistance <- st_distance(calvesSp, waterSp)
# Plot the ones that are within the SpTh
## Plot the calves 
calvesSp %>%
  filter(WaterDistance <= 0.333) %>% # Filter the ones that are below the SpTh
  st_geometry() %>% 
  plot(pch = 16, cex = 0.5, main = "Contacts within a SpTh = 0.333")
## Polot the water source:
waterSp %>%
  st_geometry() %>%
  plot(add = T)
```

## Identify the optimal SpTh value that will allow us to capture 99% of the contacts, defined as instances when point locations were within 0.33 m of the water trough
```{r}
SpThValues<-contact::findDistThresh(n1 = 1000, n2 = 1000, acc.Dist1 = 0.5, acc.Dist2 = NULL, pWithin1 = 90, pWithin2 = NULL, spTh = 0.5)
SpThValues
CI_99<-unname(SpThValues[21]) #we will use this SpTh value moving forward.
```

`SpThValues` will have the distances that captures the % of the population, we will use the value 0.7353102 which represents the 99%.  

```{r}
calvesSp %>%
  filter(WaterDistance <= CI_99) %>% # Filter the ones that are below the SpTh
  st_geometry() %>% 
  plot(pch = 16, cex = 0.5, main = "Contacts within a SpTh = 99%")
## Polot the water source:
waterSp %>%
  st_geometry() %>%
  plot(add = T)
```


```{r}
water_edges <- calvesSp %>% 
  data.frame() %>%
  filter(WaterDistance <= CI_99) %>%
  group_by(from = calftag) %>%
  summarise(to = "water", durations = n())
```

```{r}
calvesSp %>% data.frame() %>%
  filter(WaterDistance <= CI_99) %>%
  arrange(calftag, DateTime) %>% # Order by time and ID
  mutate(Interval = DateTime - lag(DateTime)) %>% # Difference from previous 
  group_by(calftag) %>%
  mutate(Interval = replace(Interval, row_number() == 1, NA)) %>%
  ungroup(calftag) %>%
  mutate(Contact = ifelse(abs(Interval) >= (5*60) | is.na(Interval), 1:n(), NA)) %>% 
  do(zoo::na.locf(.))# Identify if contact was initiated
  # group_by(calftag, Contact) %>%
  # summarise(Duration = sum(Interval, na.rm = T), DateTime = first(DateTime))
  
```

```{r}
Calves_Edges <- calvesSp %>% data.frame() %>%
  filter(WaterDistance <= CI_99) %>%
  arrange(calftag, DateTime) %>% # Order by time and ID
  mutate(Interval = DateTime - lag(DateTime)) %>% # Difference from previous 
  group_by(calftag) %>% # Gourp by calftag for identifying intervals
  mutate(Interval = replace(Interval, row_number() == 1 | abs(Interval) >= (5*60), NA)) %>% # New contacts defined for a contact with interval larger than 5 min
  ungroup(calftag) %>%
  mutate(Contact = ifelse(is.na(Interval), 1:n(), NA)) %>% # get a contact ID based on the number of row
  do(zoo::na.locf(.)) %>% # Replace NAs with the ContactID
  group_by(from = calftag, Contact) %>% # Sum the duration by contact ID
  summarise(Duration = sum(Interval), DateTime = first(DateTime)) %>%
  mutate(to = "water") %>%
  select(from, to, DateTime, Duration, Contact)
```


```{r}
G1 <- igraph::graph_from_data_frame(Calves_Edges)
```


```{r}
plot(G1)
```

We will define our distance treshold based on what we did before. Might not mkake a lot of sense.  

```{r}
# FUnction to get the contacts
CreateNetwork <- function(DF, DTh){
  # DF <- calves
  # DTh <- CI_99
  TimePoints <- DF %>% # Get unique timepoints
    distinct(DateTime)
  
  D_DF <- data.frame(Var1 = NA, Var2 = NA, Distance = NA, DateTime = TimePoints[1,]) # Create a data frame to fill with the data
  
  # Loop to get the contacts
  for(i in 1:nrow(TimePoints)){
    # i <- 1
    SP_i <- calvesSp %>% 
      filter(DateTime == TimePoints[i,])
    
    D_DF_i <- expand.grid(SP_i$calftag, SP_i$calftag) # Create a data frame for all combinations of possible Origin-Destination
    V1 <- match(D_DF_i$Var1, SP_i$calftag) # match variables from the O-D data frame to the spatial points for origins
    V2 <- match(D_DF_i$Var2, SP_i$calftag) # match variables from the O-D data frame to the spatial points for destinations
    # Calculate distance
    D_DF_i$Distance <- st_distance(SP_i[V1,], # order by matched origin variables
                                   SP_i[V2,], # order by matched Destination variables 
                                   by_element = T) 
    
    D_DF_i <- D_DF_i %>%
      mutate(DateTime = TimePoints[i,]) %>%
      filter(Var1 != Var2) %>% # Exclude the ones that are same origin-destination
      filter(Distance <= DTh) # Filter to the ones that are under the Spatial treshold
    
    D_DF <- rbind(D_DF, D_DF_i)
  }
  return(D_DF)
}



Calces_Edges <- CreateNetwork(DF = calves, DTh = 10)
```

```{r}
Calces_Edges %>%
 group_by(Distance, DateTime) %>%
  summarise(V1 = first(Var1), V2 = first(Var2)) %>%
  head(10)
```

The next ste will be to identify unique contacts like we did in the water contacts, I'll have a look to that and update the code


