---
title: "Batch Geocode"
author: "Pablo Gomez"
---

Geocoing in R can be tricky, there are many packages that can do the job, one of my favorites is `taRifx.geo` which queries trought bing maps. In order to use the service you must get a key, [Get here](https://www.bingmapsportal.com/).  
Here we will do an example of geocoding addressess in washington DC from different locations.  

First we will get a table with the name of the bulding and the address.  

```{r}
# Construct a table of addresses
dc_addresses <- data.frame(name = c("White House", 
                                    "National Academy of Sciences", 
                                    "Department of Justice", "Supreme Court", 
                                    "Washington Monument"), 
                           addr = c("1600 Pennsylvania Ave Washington, DC", 
                                    "2101 Constitution Ave NW, Washington, DC 20418", 
                                    "950 Pennsylvania Ave NW, Washington, DC 20530", 
                                    "1 1st St NE, Washington, DC 20543", 
                                    "2 15th St NW, Washington, DC 20024"))
```

Next we will load the library and we put our Bing maps key
```{r}
library(taRifx.geo)
options(BingMapsKey = "AtTAWG0w8AC5alcTk2swTE8R24GuTDRoZWTv75gizt-1FBULSUaVtMDQ3Y35ColY")
```

The finction `geocode` will give us what we want, we use verbose to get a confidence in the geocoding query, the confidence can be high, medium or low.  
```{r}
geocode(dc_addresses$addr[1], service = "bing", verbose = T)
```

This fucntion can only do one address at the time, if we want to geocode multiple addresses we will need to create a function:

```{r}
batch_geocode <- function(Address){
  Lat <- Lon <- vector()
  for(i in 1:length(Address)){
    GC <- geocode(Address[i], service = "bing")
    Lat <- c(Lat, GC[1])
    Lon <- c(Lon, GC[2])
  }
  return(data.frame(Lat, Lon))
}


# Now we will bind our geocoded columns to the data:
geocoded_addresses <- cbind(dc_addresses, 
                            batch_geocode(Address = dc_addresses$addr))
```

We can then convert the coordinates to a spatial object:

```{r}
library(sf)
Pts <- geocoded_addresses %>%
  st_as_sf(coords = c("Lon", "Lat"), crs = st_crs(4326))
```

Then we use tmap to plot our locations in an interactive view.
```{r}
library(tmap)
tmap_mode("view")
tm_shape(Pts) +
  tm_dots(col = "red")
```
